---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Linear Mixed Effects Models}
   \rfoot{\color{headergrey}Chapter 7}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Bayesian Data Analysis}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(cowplot, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(lattice, quietly = TRUE, warn.conflicts = FALSE)

library(blmeco, quietly = TRUE, warn.conflicts = FALSE)
library(tidybayes, quietly = TRUE, warn.conflicts = FALSE)
library(arm, quietly = TRUE, warn.conflicts = FALSE)
library(coda, quietly = TRUE, warn.conflicts = FALSE)
library(nlme, quietly = TRUE, warn.conflicts = FALSE)
library(sp, quietly = TRUE, warn.conflicts = FALSE)
library(gstat, quietly = TRUE, warn.conflicts = FALSE)
library(lme4, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Linear Mixed Effects Models

## Why Mixed Effects Models?

### Pooling

Complete:

$\hat{y}_i = \beta_o$
$y_i \sim Norm(\hat{y}_i, \sigma^2)$

Partial:

$\hat{y_i} = \beta_0 + b_{g_{i}}$
$y_i \sim Norm(\hat{y_i}, \sigma^2)$
$b_g \sim Norm(0, \sigma^2_b)$

No pooling:

$\hat{y} = \beta_{g_i}$
$y_i \sim Norm(\hat{y_i}, \sigma^2)$

## Fitting Linear Mixed Models

```{r}
data("cortbowl")

dat <- cortbowl
mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Ring), 
            data = dat, REML = T)

summary(mod)

round(fixef(mod), 3)

ranef(mod)
```

## Restricted Maximum Liklihood Estimation

```{r}
mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Ring), 
            data = dat, REML = F) # use ML

```

## Assessing Model Assumptions

```{r}
par(mfrow=c(2,2))
scatter.smooth(fitted(mod), resid(mod)); abline(h=0, lty=2)
title("Tukey-Anscombe Plot")

qqnorm(resid(mod), main = "norm QQ-plot, residuals") # qq of residuals
qqline(resid(mod))
scatter.smooth(fitted(mod), sqrt(abs(resid(mod))))

qqnorm(ranef(mod)$Ring[, 1], main = "normal QQ-plot, random effects")
qqline(ranef(mod)$Ring[, 1]) # qq random effects
```

## Drawing Conclusions

```{r}
nsim <- 2000
bsim <- sim(mod, n.sim = nsim)
str(bsim)

round(apply(bsim@fixef, 2, quantile, prob=c(0.025, 0.5, 0.975)), 3)

```

```{r}

imp_factor <- factor(c("C", "P"), levels = levels(dat$Implant))
days_factor <- factor(c(1, 2, 3))

newdat <- expand.grid(Implant = imp_factor, days = days_factor)

Xmat <- model.matrix(~Implant + days + Implant:days, data = newdat)
fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))
for(i in 1:nsim) fitmat[, i] <- Xmat %*% bsim@fixef[i, ] # fitted values
newdat$lower <- apply(fitmat, 1, quantile, prob = 0.025)
newdat$upper <- apply(fitmat, 1, quantile, prob = 0.975)
newdat$fit <- Xmat %*% fixef(mod)

ggplot(newdat, aes(days, fit, color = Implant)) +
   geom_point()

```

## Frequentist Results

```{r}

```

## Random Intercept and Random Slope

```{r}

data(wingbowl)
dat <- wingbowl
dat$Age.z <- scale(dat$Age)
mod <- lmer(Wing ~ Age.z + Implant + Age.z:Implant + (Age.z|Ring),
          data = dat, REML = F)
mod
```

```{r}

par(mfrow=c(1, 2))
qqnorm(ranef(mod)$Ring[, 1]) # Intercept
qqline(ranef(mod)$Ring[, 1])

qqnorm(ranef(mod)$Ring[, 2]) # Slope
qqline(ranef(mod)$Ring[, 2])

nsim <- 2000
bsim <- sim(mod, n.sim = nsim)
apply(bsim@fixef, 2, quantile, prob = c(0.025, 0.975))

quantile(bsim@fixef[, "Age.z:ImplantP"] / sd(dat$Age), prob = c(0.025, 0.975))
```

```{r}
newdat <- expand.grid(Age = seq(23, 45, length = 100),
                      Implant = levels(dat$Implant))

newdat$Age.z <- (newdat$Age - mean(dat$Age)) / sd(dat$Age)

head(newdat)
dim(newdat)

Xmat <- model.matrix(~Age.z + Implant + Age.z:Implant, data = newdat)
fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))

dim(fitmat)
dim(bsim@fixef)

head(bsim@fixef)

for(i in 1:nsim) {
   fitmat[, i] <- Xmat %*% bsim@fixef[i, ]
}

dim(bsim@fixef)

newdat$lower <- apply(fitmat, 1, quantile, prob = 0.025)
newdat$upper <- apply(fitmat, 1, quantile, prob = 0.975)
```

```{r}
par(mfrow = c(1, 2), mar=c(5, 1, 1, 1), oma = c(0, 4, 0, 0))
plot(dat$Age.z, dat$Wing, pch=1, cex = 0.8, las = 1,
     col = c("orange", "blue")[as.numeric(dat$Implant)],
     xlab = "Age (days)", ylab = NA, xaxt = "n")
at.x_orig <- seq(25, 45, by = 5) # values on the x-axis, original scale
at.x <- (at.x_orig - mean(dat$Age)) / sd(dat$Age) # transformed scale
axis(1, at = at.x, labels = at.x_orig) # original values at transformed
mtext("Wing length (mm)", side = 2, outer = T, line = 2, cex = 1.2, adj=0.6)
abline(fixef(mod)[1], fixef(mod)[2], col = "orange", lwd=2) # for C
abline(fixef(mod)[1] + fixef(mod)[3], fixef(mod)[2] + fixef(mod)[4],
       col = "blue", lwd = 2)

for(i in 1:2) {
   index <- newdat$Implant == levels(newdat$Implant)[i]
   
   x <- c(newdat$Age.z[index], rev(newdat$Age.z[index]))
   y <- c(newdat$lower[index], rev(newdat$upper[index]))
   
   polygon(x, y,
           border = NA, col = c(rgb(1, 0.65, 0, 0.5), rgb(0, 0, 1, 0.5))[i])
}

plot(dat$Age.z, dat$Wing, pch=1, cex=0.8, las=1,
     col = c("orange", "blue")[as.numeric(dat$Implant)],
     xlab = "Age (days)", ylab = NA, yaxt = "n", xaxt = "n")
at.x_orig <- seq(25, 45, by = 5)
ax.x <- (at.x_orig - mean(dat$Age)) / sd(dat$Age)
axis(1, at = at.x, labels = at.x_orig)
indtreat <- tapply(dat$Implant, dat$Ring, function(x) as.character(x[1]))
for(i in 1:86) {
   if(indtreat[i] == "C") abline(fixef(mod)[1] + ranef(mod)$Ring[i, 1],
                                 fixef(mod)[2] + ranef(mod)$Ring[i, 2],
                                 col = "Orange") else
                                    abline(fixef(mod)[1] + fixef(mod)[3],
                                           ranef(mod)$Ring[i, 1], fixef(mod)[2] +
                                              fixef(mod)[4] + ranef(mod)$Ring[i, 2],
                                           col = "blue")
}

```

## Nested and Crossed Random Effects

```{r}
data(cortbowl)
dat <- cortbowl

mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Brood) + (1|Ring),
            data = dat, REML = F)

mod

```

```{r}
mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Brood/Ring),
            data = dat, REML = F)
mod
```

```{r}
data("ellenberg")

ellenberg$gradient <- paste(ellenberg$Year, ellenberg$Soil)
table(ellenberg$Species, ellenberg$gradient)
```

```{r}
ellenberg$Water.z <- as.numeric(scale(ellenberg$Water))
mod <- lmer(log(Yi.g) ~ Water.z + I(Water.z^2) +
               (Water.z + I(Water.z^2)|Species) + (1|gradient),
            data = ellenberg)

mod
```

```{r}

```

