---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 8}
   \fancyfoot[C]{\rmfamily\color{headergrey}Tree Based Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(tree, quietly = TRUE, warn.conflicts = FALSE)
library(randomForest, quietly = TRUE, warn.conflicts = FALSE)
library(gbm, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 8

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Lab

### Fitting Classification Trees


```{r}
carseats <- as.data.table(ISLR::Carseats)

carseats[, High := as.factor(ifelse(Sales <= 8, "No", "Yes"))]

tree.carseats <- tree(formula = High ~ .-Sales, data = carseats)

summary(tree.carseats)

plot(tree.carseats)
text(tree.carseats, pretty = 0)
```

```{r}
set.seed(2)

train <- sample(1:nrow(carseats), 200)

carseats.test <- carseats[-train]
high.test <- carseats[-train]$High

tree.carseats <- tree(High ~.-Sales, data = carseats, subset = train)
tree.pred <- predict(tree.carseats, carseats.test, type = "class")

table(tree.pred, high.test)
```

```{r}
set.seed(3)

cv.carseats <- cv.tree(tree.carseats, FUN = prune.misclass)
names(cv.carseats)

cv.carseats
```

```{r}
par(mfrow = c(1, 2))
plot(cv.carseats$size, cv.carseats$dev, type = "b")
plot(cv.carseats$k, cv.carseats$dev, type="b")
```

```{r}
par(mfrow = c(1,1))
prune.carseats <- prune.misclass(tree.carseats, best = 9)
plot(prune.carseats)
text(prune.carseats, pretty = 0)
```

```{r}
tree.pred <- predict(prune.carseats, carseats.test, type = "class")
table(tree.pred, high.test)
```

```{r}
prune.carseats <- prune.misclass(tree.carseats, best = 15)
plot(prune.carseats)
text(prune.carseats, pretty = 0)

tree.pred <- predict(prune.carseats, carseats.test, type = "class")
table(tree.pred, high.test)
```

### Regression Trees

```{r}

boston <- as.data.table(Boston)

N <- nrow(boston)

set.seed(1)

train <- sample(1:N, N /2)

tree.boston <- tree(medv ~ ., boston, subset = train)

summary(tree.boston)
```

```{r}
plot(tree.boston)
text(tree.boston, pretty = 0)
```

```{r}
cv.boston <- cv.tree(tree.boston)
plot(cv.boston$size, cv.boston$dev, type = 'b')
```

```{r}
prune.boston <- prune.tree(tree.boston, best = 5)
plot(prune.boston)
text(prune.boston, pretty = 0)
```

```{r}
yhat <- predict(tree.boston, newdata = boston[-train,])
boston.test <- boston[-train]$medv

plot(yhat, boston.test)
abline(0, 1)
mean((yhat - boston.test)^2)
```

### Bagging and Boosting

```{r}
set.seed(1)

bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, importance = T)
bag.boston
```

```{r}
boston.test <- boston[!train]
yhat.bag <- predict(bag.boston, newdata = boston.test)
plot(yhat.bag, boston.test$medv)
abline(0, 1)
mean((yhat.bag-boston.test$medv)^2)
```

```{r}
bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, ntree = 25)
yhat.bag <- predict(bag.boston, newdata = boston[-train,])
mean((yhat.bag - boston.test$medv)^2)
```

```{r}
set.seed(1)

rf.boston <- randomForest(medv ~  ., data = boston, subset = train, mtry = 6, importance = T)
yhat.rf <- predict(rf.boston, newdata = boston[-train,])
mean((yhat.rf-boston.test)^2)
```

```{r}
importance(rf.boston)
```

```{r}
varImpPlot(rf.boston)
```

### Boosting

```{r}
boost.boston <- gbm(medv ~ ., data = boston[-train,], distribution = "gaussian", n.trees = 5000, interaction.depth = 4)

summary(boost.boston)
```

```{r}
par(mfrow = c(1, 2))
plot(boost.boston, i = "rm")
plot(boost.boston, i = "lstat")
```

```{r}
yhat.boost <- predict(boost.boston, newdata = boston[-train,], n.trees = 5000)
mean((yhat.boost - boston.test)^2)
```

## Applied

In the lab, we applied random forests to the *Boston* data using mtry = 6 and using ntree = 25 and ntree = 500. Create a plot displaying the test error resulting from random forests on this data set for a more comprehensive range of values for mtry ntree.

```{r}
N <- nrow(boston)

train <- sample(1:N, N * .7)

boston.train <- boston[train]
boston.test <- boston[!train]

ntrees <- seq(0, 500, 10)
n.features <- ncol(boston[, !"medv"])

mtry <- round(c(n.features, n.features/2, sqrt(n.features)))

results <- matrix(vector(mode = "numeric", length(ntrees) * length(mtry)), nrow = 3)

for(i in 1:length(ntrees))
{
   for(j in 1:length(mtry))
   {
      tree.count <- ntrees[i]
      
      rf <- randomForest(medv ~ ., data = boston.train, mtry = mtry[j], n.trees = tree.count)
      
      pred <- predict(rf, newdata = boston.test)
      
      results[j, i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
   }
}

df_results <- as.data.table(results)
colnames(df_results) <- paste0(ntrees)

df_results <- melt(cbind(mtry, df_results), id.vars = "mtry")
df_results$mtry <- as.factor(df_results$mtry)

ggplot(df_results, aes(variable, value, group = mtry)) +
   geom_line(aes(col = mtry))

```

### Carseats

In the lab, a classification tree was applied to the Carseats data set after converting Sales into a qualitative response variable. Now we will seek to predict Sales using regression trees and related approaches, treating the response as a quantitaive variable.

a.) Split the data into a t training set and a test set.

```{r}
N <- nrow(carseats)

index <- sample(1:N, N * .7)

carseats.train <- carseats[index]
carseats.test <- carseats[!index]
```

b.) Fit a regression tree to the training set. Plot the tree, and interpret the results. What test MSE do you obtain?

```{r}
carseat.tree <- tree(Sales ~., data = carseats.train)
pred <- predict(carseat.tree, newdata = carseats.test)
mse <- mean((pred - carseats.test$Sales)^2)

mse

par(mfrow = c(1,1))
plot(carseat.tree)
text(carseat.tree, pretty = 0)
```

c.) Use cross-validation in order to determine the optimal level of tree complexity. Does pruning the tree improve the test MSE?

```{r}
carseat.cv <- cv.tree(carseat.tree)

plot(carseat.cv$size, carseat.cv$dev, type = "b", col = "red")
plot(carseat.cv$size, carseat.cv$k, type = "b", col = "red")

prune.carseats <- prune.tree(carseat.tree, best = 5)

plot(prune.carseats)
text(prune.carseats)
```

d.) Use the bagging approach in order to analyze this data. What test MSE do you obtain? Use the importance() function to determine which variables are most important.

```{r}
p <- ncol(carseats)

ntrees <- seq(0, 500, 25)
rf.error <- numeric(length(ntrees))

for(i in 1:length(ntrees))
{
   carseats.rf <- randomForest(Sales ~ ., data = carseats.train, mtry = p, n.trees = ntrees[i])
   
   pred <- predict(carseats.rf, newdata = carseats.test)
   
   rf.error[i] <- mean((pred - carseats.test$Sales)^2)
}

lowest.error <- which.min(rf.error)

results <- data.table(trees = ntrees, error = rf.error)[, lowest := .I == lowest.error]

ggplot(results, aes(trees, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = paste0("lowest error: #", ntrees[lowest.error]))

```

e.) Use random forest to analyize this data. What test MSE do you obtain? Use the importance() fucntion to determine which variables are most important. Describe the effect of m, the number of variables considered at each split, on the error obtained.

```{r}
p <- ncol(carseats)

mtry <- seq(1, p)

rf.error <- numeric(p)

for(i in 1:p)
{
   carseats.rf <- randomForest(Sales ~ ., data = carseats.train, mtry = mtry[i], n.trees = 250)
   
   pred <- predict(carseats.rf, newdata = carseats.test)
   
   rf.error[i] <- mean((pred - carseats.test$Sales)^2)
}

lowest.error <- which.min(rf.error)

results <- data.table(mtry = mtry, error = rf.error, lowest = mtry == lowest.error)

ggplot(results, aes(mtry, error, fill = lowest)) +
   geom_bar(stat = "identity")
```

### OJ

This problem involves the **OJ** data set which is part of the ISLR package.

a.) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.

```{r}
oj <- as.data.table(ISLR::OJ)
n <- nrow(oj)

index <- sample(1:n, n *.7, replace = F)

oj.train <- oj[index]
oj.test <- oj[!index]

head(train)
```

b.) Fit a tree to the training data, which **Purchases** as the first response and the other variables as predictors. Use the **summary** function to produce summary statistics about the tree, and describe the results obtained.

```{r}

oj.tree <- tree(Purchase ~., data = oj.train)

plot(oj.tree)
text(oj.tree)

summary(oj.tree)

```

What is the training error rate? _0.7469_

How many terminal nodes? _8_

c.) Type in the name of the tree object in order to get a detailed text output. Pick one of the terminal nodes, and interpret the information displayed.

```{r}
summary(oj.tree)
```

d.) Create a plot of the tree.

```{r}
plot(oj.tree)
text(oj.tree)
```

e.) Predict the response on the test data, and produce a confusion matrix comparing the test labels to the predicted test labels.

```{r}
pred <- predict(oj.tree, newdata = oj.test, type = "class")

confusionMatrix(oj.test$Purchase, pred)
```

f.) Apply the cv.tree function to the training data set in order to determine the optimal tree size.

```{r}
oj.cv <- cv.tree(oj.tree)
```

g.) Produce a plot with the tree size and cross-validated error rate.

```{r}
plot(oj.cv$size, oj.cv$dev, type ='b', col =  "cornflowerblue") # 5
```

h.) What tree size corresponds to the lowest cv error rate?

**5**

i.) Produce a pruned tree corresponding to the optimal tree size obtained using cross-validation.

```{r}
oj.pruned <- prune.tree(oj.tree, k = 5)
```

j.) Compare the training error rates between the pruned and unpruned trees.

```{r}

summary(oj.tree)
summary(oj.pruned)
```

Same.

k.) Compare the test error rates.

```{r}
pruned.pred <- predict(oj.pruned, newdata = oj.test, type = "class")

confusionMatrix(oj.test$Purchase, pred)

confusionMatrix(oj.test$Purchase, pruned.pred)
```

### Hitters

We now use boosting to predict **Salary** in the hitters data set.

a.) Remove the observations for whom the salary information is unknown, and then log-transform the salaries.

```{r}
hitters <- as.data.table(ISLR::Hitters)
hitters <- hitters[complete.cases(hitters)]
hitters$Salary <- log(hitters$Salary)

```

b.) Create a training set consisting of the first 200 observations and a test set containing the rest.

```{r}
n <- nrow(hitters)

index <- 1:200

hitters.train <- hitters[index]
hitters.test <- hitters[!index]
```

c.) Perform boosting on the training set with 1,000 trees for a range of values of the shrinkage parameter $\lambda$. Produce a plot with the different shrinkage values and the training MSE.

```{r}
shrinkage <- seq(0.001, 0.5, by = 0.0025)
iters <- length(shrinkage)

results <- vector(mode = "numeric", length = iters)

for(i in 1:iters)
{
   gbm <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[i], distribution = "gaussian")
   results[i] <- mean(gbm$train.error^2)
}

lowest.error <- which.min(results)

df_results <- data.table(shrinkage, error = results)[, lowest := .I == lowest.error]

ggplot(df_results, aes(shrinkage, error, fill = lowest)) +
   geom_bar(stat = "identity")
```

d.) Produce a plot with different shrinkage methods and the corresponding test mse.

```{r}
shrinkage <- seq(0.001, 0.5, by = 0.0025)
iters <- length(shrinkage)

results <- vector(mode = "numeric", length = iters)

for(i in 1:iters)
{
   gbm <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[i], distribution = "gaussian")
   
   pred <- predict(gbm, n.trees = 1000, newdata = hitters.test, type = "response")
   
   results[i] <- mean((hitters.test$Salary - pred)^2)
}

lowest.error <- which.min(results)

df_results <- data.table(shrinkage, error = results)[, lowest := .I == lowest.error]

ggplot(df_results, aes(shrinkage, error, fill = lowest)) +
   geom_bar(stat = "identity")

tree.mse <- results[lowest.error]
```

e.) Compare the test MSE of boosting to the test MSE that results from applying two of the regression approaches seen in chapters 3 & 6.

```{r}

train.mat <- model.matrix(Salary ~., data = hitters.train)
test.mat <- model.matrix(Salary ~., data = hitters.test)

cv.ridge <- cv.glmnet(train.mat, hitters.train$Salary, data = hitters.train, alpha = 0)

ridge.fit <- glmnet(test.mat, hitters.test$Salary, alpha = 1, lambda = cv.ridge$lambda.min)
ridge.pred <- predict(ridge.fit, newx = test.mat, type = "response")
ridge.error <- mean((hitters.test$Salary - ridge.pred)^2)

cv.lasso <- cv.glmnet(train.mat, hitters.train$Salary, data = hitters.train, alpha = 1)

lasso.fit <- glmnet(test.mat, hitters.test$Salary, alpha = 0, lambda = cv.lasso$lambda.min)
lasso.pred <- predict(lasso.fit, newx = test.mat, type = "response")
lasso.error <- mean((hitters.test$Salary - lasso.pred)^2)

results <- data.table(type = c("Tree", "Ridge", "Lasso"), error = c(tree.mse, ridge.error, lasso.error))

ggplot(results, aes(type, error, fill = type)) +
   geom_bar(stat = "identity")
```

f.) Which variables appear to be the most important predictors in the boosted model?

```{r}

best.tree <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[lowest.error], distribution = "gaussian")

vip::vip(best.tree)
```

### Caravan

This question uses the caravan data set.

a.) Create a training set consisting of the first 1,000 observations, and a test set consisting of the remaining observations.

```{r}
caravan <- as.data.table(ISLR::Caravan)
caravan$Purchase <- ifelse(caravan$Purchase == "Yes", 1, 0)
index <- 1:1000

caravan.train <- caravan[index]
caravan.test <- caravan[-index]

head(caravan.train)
```

b.) Fit a boosting model to the training set with **Purchases** as the response and the other variables as predictors. Use 1,000 trees and a shrinkage value of .01.

```{r}
boost.fit <- gbm(Purchase ~ ., data = caravan.train, n.trees = 1000, shrinkage = 0.1, distribution = "bernoulli")
```

What predictors appear to be most important?

```{r}
vip::vip(boost.fit)
```

c.) Use the boosting model to predict the response on the test data. Predict that a person will make a purchase if the estimated probablity of purchase is greater than 20%.

```{r}
probs <- exp(predict.gbm(boost.fit, newdata = caravan.test, type = "response", n.trees = 1000)) - 1

pred <- ifelse(probs > .2, 1, 0)

confusionMatrix(as.factor(caravan.test$Purchase), as.factor(pred))
```

