---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 1}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 2

### 1.1

Store the values _-20, -15, -5, 8, 12, 9, 2, 23, 19_ in the R variable _x_ and verify the sum is 33.

```{r, echo = T}
x <- c(-20, -15, -5, 8, 12, 9, 2, 23, 19)

sum(x)
```

### 1.2

For the data in E1, verify the average is 3.67.

```{r, echo = T}
mean(x)
```

### 1.3

What commands can be used to compute an average without using the R command _mean_?

```{r, echo = T}
n <- length(x)

sum(x) / n

```

### 1.4

Sum the positive values:

```{r, echo = T}
sum( x[x >= 0] )
```

### 1.5

Use _which_ command to get the average of the values ignoring the largest.

```{r, echo = T}

mean( which( x != max(x) ) )

```

### 1.6

Speculate about the values corresponding to _x[ abs(x) >= 8 & x < 8 ]_.

[T, T, F, T, T, F, T, T]
&
[T, T, T, F, F, F, F, F]
=
[T, T, F, F, F, F, F, F]

Verify.

```{r, echo = T}
x[ abs(x) >= 8 & x < 8]
```

### 1.7

Your recorded time to commute for 10 days is:

23, 18, 29, 22, 24, 27, 28, 19, 28, 23

```{r, echo = T}
commute <- c(23, 18, 29, 22, 24, 27, 28, 19, 28, 23)

mean(commute)
min(commute)
max(commute)

```

### 1.8

Verify:

```{r, echo = T}
y <- c(2, 4, 8)
z <- c(1, 5, 2)
2*y
```

returns the values _4, 8, 16_.

### 1.9

Let _x = c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)_.

```{r, echo = T}
x <- c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)

n <- length(x)

stopifnot( (sum(x) / n) == mean(x))

```

### 1.10

For the values in _E9_, use R to subtract the average from each value, then sum the results.

```{r, echo = T}
xbar <- mean(x)
sum(x - xbar)
```

### 1.11

Imagine a matrix having 100 rows and 2 columns. Further imagine that some of the values in the first column are _NA_.

Use _is.na_ to remove missing.

```{r, echo = T}

values <- matrix(nrow = 100, ncol = 2)
rval <- runif(100)
values[, 1] <- ifelse(rval <= .25, NA, rval)


values[!is.na(values[,1]),]

```

### 1.12

R has a built-in data set ChickWeight. Verify the mean.

```{r, echo = T}
mean(ChickWeight[, 1])

mean(ChickWeight[, 3])

mean(as.numeric(ChickWeight[, 3]))

```

### 1.13

Create a matrix with two rows and five columns with some of the entries stored as NA.


```{r, echo = T}

rval <- runif(10, 0, 1)
values <- matrix(ifelse(rval <= .25, NA, rval), nrow = 2)

```

### 1.14

Use R to compute the average weight among chicks that were fed horsebean.

```{r, echo = T}
mean(chickwts[chickwts$feed == "horsebean",]$weight)
```

### 1.15

Let _x = 1, 8, 2, 6, 3, 8, 5, 5, 5, 5_

Sum values not equal to 2 or 3, two ways.

```{r, echo = T}
x <- c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)

sum( x[!x %in% c(2, 3)] )

sum( x[x != 2 & x != 3] )

```

### 1.16

For the values used in the previous exercise, use two different R commands to sum all the values != 5.

```{r, echo = T}
sum( x[ x != 5] )
sum( x[ !x %in% c(5) ])
```

### 1.17

Use a single command to change all values equal to 8 to 7.

```{r, echo = T}
x[x == 8] <- 7
x  
```

### 1.18

Create a matrix with four rows and two columns with the values: _1, 2, 3, 4_ in the first column, and _5, 6, 7, 8_ in the second column.

```{r, echo = T}
matrix(seq(1, 8, by = 1), nrow = 4, ncol = 2)
```

### 1.19

Create a matrix with four columns and two rows with the values _1, 2, 3, 4_ in the first row and _11, 12, 13, 14_ in the second row.

```{r, echo = T}
matrix(c(seq(1,4), seq(11, 14)), ncol = 4, byrow = T)
```