---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Numerical Summaries}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 2}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 2

### 2.1

Suppose,

$X_1 = 1, X_2 = 3, X_3 = 0,$
$X_4 = -2, X_5 = 4, X_6 = -1,$
$X_7 = 5, X_8 = 2, X_9 = 10.$

```{r, echo = T}
x <- c(1, 3, 0, -2, 4, -1, 5, 2, 10)
```

Find:

a.) $\sum{X_i}$

```{r, echo = T}
sum(x)
```

b.) $\sum^5_{i=3}X_i$

```{r, echo = T}
sum( x[3:5] )
```

c.) $\sum^4_{i=1}X^3_i$

```{r, echo = T}
sum(x[1:4]^3)
```

d.) $(\sum{X_i})^2$

```{r, echo = T}
(sum(x))^2
```

e.) $\sum{3}$

```{r, echo = T}
3 * length(x)
```

f.) $\sum{(X_i - 7)}$

```{r, echo = T}
sum(x - 7)
```

g.) $3\sum^5_{i=1}X_i - \sum^9_{i=6}X_i$

```{r, echo = T}
3 * sum( x[1:5] ) - sum( x[6:9] )
```

h.) $\sum{10X}$

```{r, echo = T}
sum( 10 * x)
```

i.) $\sum^6_{i=2}iX_i$

```{r, echo = T}
i <- 2:6
sum( i * x[i]  )
```

j.) $\sum6$

```{r, echo = T}
6 * length(x)
```

### 2.2

Express the following in summation notation.

a.) $X_1 + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$

$\ldots = \frac{X_1}{1} + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$

$\ldots = \sum{^4}_{i=1}{\frac{X_i}{i}}$

b.) $U_1 + U^2_2 + U^3_3 + U^4_4$

$\ldots = U^1_1 + U^2_2 + U^3_3 + U^4_4$

$\ldots = \sum^4_{i=1}{U^i_i}$

c.) $(Y_1 + Y_2 + Y_3)^4$

$(\sum^3_{i=1}{Y_i})^4$

### 2.3

Show by numerical example that $\sum{X^2_i}$ is not necessarily equal to $(\sum{X_i}^2)$.

```{r, echo = T}
x <- 1:10

sum(x^2)

(sum(x))^2
```

### 2.4

Find the mean and median of the following sets of numbers.

a.) -1, 0, 3, 0, 2, -5.

```{r, echo = T}
x <- c(-1, 0, 3, 0, 2, -5)

mean(x)
median(x)
```

b.) 2, 2, 3, 10, 100, 1,000

```{r, echo = T}
x <- c(2, 2, 3, 10, 100, 1000)

mean(x)
median(x)
```

### 2.5

The final exam scores for 15 students are: 73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99.

Compute the mean, 20% trimmed mean, and median using R.

```{r, echo = T}
scores <- c(73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99)

mean(scores)
mean(scores, trim = .2)
median(scores)
```

### 2.6

The average of 23 numbers is 14.7. What is the sum of these numbers?

```{r, echo = T}
23 * 14.7
```

### 2.7

Consider the 10 values: 3, 6, 8, 12, 23, 26, 37, 42, 49, 63.

The mean is $\bar{X} = 26.9$

```{r, echo = T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 63)
mean(x)
```

a.) What is the value of the mean if the largest value, 63, is increased to 100?

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 100)
mean(x)
```

b.) What is the mean if 633 is increased to 1,000?

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 1000)
mean(x)
```

c.) What do these results illustrate about the mean?

_The mean is very sensitive to outliers._

### 2.8

Repeat the previous exercise, only compute the median instead.

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 100)
median(x)
```

b.) What is the mean if 633 is increased to 1,000?

```{r, echo=T}
x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 1000)
median(x)
```

### 2.9

In general, how many values must be altered to make the sample mean arbitrarily large?

_One._

### 2.10

What is the minimum number of values that must be altered to make the 20% trimmed mean and sample median arbitrarily large?

mean = $g = (.2N), g + 1$

median = ~$.5N$