---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Confidence Intervals}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(here, quietly = T, warn.conflicts = F)
library(ggplot2, quietly = T, warn.conflicts = F)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 6

### 6.1

Explain the meaning of a .95 confidence interval.

_A confidence interval is a statistical measure of the probability coverage a given lies within some interval based upon the observed data._

### 6.2

If the goal is to compute a .80, or .92, or a .98 confidence interval for $\mu$ when $\sigma$ is known, and sampling is from a normal distribution, what values for c should be used in Equation (6.4)?

```{r, echo = T}
pretty_kable(
   data.table( 
      Confidence = c(.8, .92, .98) )[, 
         Val := (1 - (1 - Confidence)/2)][, C := qnorm(Val)] , 
   "Confidence Values")
```

### 6.3

```{r, echo = T}
conf <- function(alpha) {
   c( Lower = (1 - alpha)/2, Upper = 1 - (1 - alpha)/2)   
}
```

Assuming random sampling is from a normal distribution with standard deviation $\sigma = 5$, if the sample mean is $\bar{X} = 45$ based on $n = 25$ participants, what is the 0.95 confidence interval for $\mu$?

```{r, echo = T}
qnorm(conf(.95), mean = 45, sd = 5/sqrt(25))
```

### 6.4

Repeat the previous example, only compute a .99 confidence interval instead.

```{r, echo = T}
qnorm(conf(.99), mean = 45, sd = 5/sqrt(25))
```

### 6.5

A manufacturer claims that their light bulbs have an average life span that follows a normal distribution with $\mu = 1,200$ hours and a standard deviation of $\sigma = 25$. If you randomly test 36 light bulbs and find that their average life span is $\bar{X} = 1,150$, does a .95 confidence interval for $mu$ suggest that the claim $mu = 1,200$ is reasonable?

```{r, echo = T}
qnorm(conf(.95), mean = 1150, sd = 25/sqrt(36))
```

No, 1,200 is outside the bounds of a 95% confidence interval.

### 6.6

Compute a .95 confidence interval for the mean, assuming normality, for the following situations:

a.) $n = 12, \sigma = 22, \bar{X} = 65$

```{r, echo = T}
qnorm(conf(.95), mean = 65, sd = 22/sqrt(12))
```

b.) $n = 22, \sigma = 10, \bar{X} = 185$

```{r, echo = T}
qnorm(conf(.95), mean = 185, sd = 10/sqrt(22))
```

c.) $n = 50, \sigma = 30, \bar{X} = 19$

```{r, echo = T}
qnorm(conf(.95), mean = 19, sd = 30/sqrt(50))
```

\newpage

### 6.7

What happens to the length of a confidence interval for the mean of a normal distribution when the sample size is doubled? In particular, what is the ratio of the lengths?

What is the ratio of the lengths if the sample size is quadrupled?

__Answer__:

In general, For some _x_,

$CI = \mu \pm C\frac{\sigma}{x\sqrt{n}}$

As _x_ increases, the size of the confidence intervals decrease (closer approximations).

### 6.8

The length of a bolt made by a machine parts company has a normal distribution with standard deviation $\sigma$ equal to 0.01 mm. The lengths of four randomly selected bolts are as follows:

20.01, 19.88, 20.00, 19.99

a.) Compute a 0.95 confidence interval for the mean.

```{r, echo = T}
x <- c(20.01, 19.88, 20.00, 19.99)
n <- length(x)
xbar <- mean(x)

qnorm(conf(.95), mean = xbar, sd = 0.01/sqrt(n))
```

b.) Specifications requre a mean lengths $\mu$ of 20.00 mm for the population of bolts. Do the data indicate that this specification is being met?

_No, the 95% confidence interval for the mean does not include 20._

c.) Given that the 0.95 confidence interval contains the value 20, why might it be inappropriate to conclude that the specification is being met?

_Just because the confidence interval includes some value, that does not mean that the population paramter is that value._

### 6.9

The weight of trout sold at a trout farm has a standard deviation of 0.25. Based on a sample of 10 trout, the average weight is 2.10 lbs.

Assume normality and Compute a .99 confidence interval for the mean.

```{r, echo = T}
qnorm(conf(.99), mean = 2.10, sd = .25/sqrt(10))
```

### 6.10

The average bounce of 45 randomly selected tennis balls is found to be $\bar{X} = 1.70$.

Assuming that the standard deviation of the bounce is .30, compute a 0.90 confidence interval for the average bounce assuming normality.

```{r, echo = T}
qnorm(conf(.90), mean = 1.70, sd = .30/sqrt(45))
```

### 6.11

Assuming that the degrees of freedom are 20, find the value t for which:

a.) $P(T \le t) = 0.995$

```{r, echo = T}
qt(.995, df = 20)
```

b.) $P(T \ge t) = 0.025$

```{r, echo = T}
1 - qt(.025, df = 20)
```

c.) $P(-t \le T \le t) = 0.90$

```{r, echo = T}
qt( (1 + .90) / 2, df = 20)
```

### 6.12

Compute a 0.95 confidence interval if

a.) $n = 10, \bar{X} = 26, s = 9$

```{r, echo = T}
26 + qt(conf(.95), df = 9) * 9/sqrt(10)
```

b.) $n = 18, \bar{X} = 132, s = 20$

```{r, echo = T}
132 + qt(conf(.95), df = 17) * 20/sqrt(18)
```

c.) $n = 25, \bar{X} = 52, s = 12$

```{r, echo = T}
52 + qt(conf(.95), df = 24) * 12/sqrt(25)
```

### 6.13

Compute a 0.99 confidence interval if

a.) $n = 10, \bar{X} = 26, s = 9$

```{r, echo = T}
26 + qt(conf(.99), df = 9) * 9/sqrt(10)
```

b.) $n = 18, \bar{X} = 132, s = 20$

```{r, echo = T}
132 + qt(conf(.99), df = 17) * 20/sqrt(18)
```

c.) $n = 25, \bar{X} = 52, s = 12$

```{r, echo = T}
52 + qt(conf(.99), df = 24) * 12/sqrt(25)
```

### 6.14

For a study on self-awareness, the observed values for one of the groups were:

77, 87, 88, 114, 151, 210, 219, 246, 253, 262, 296, 299, 306, 376, 428, 515, 666, 1310, 2611.

Compute a .95 confidence interval for the mean assuming normality.

```{r, echo = T}
x <- c(77,87,88,114,151,210,219,246,253,262,296,299,306,376,428,515, 666,1310,2611)
xbar <- mean(x)
n <- length(x)

xbar + qt(conf(.95), df = n - 1) * sd(x)/sqrt(n)
```

Compare the result to the bootstrap-t confidence interval.

```{r, echo = T}

winsorize <- function( x, trim = .2) {
   n <- length(x)
   
   o <- sort(x)
   g <- floor(trim*n)
   
   o[1:(g+1)] <- o[(g+1)]
   o[(n-g):n] <- o[n-g]
   
   o
}

trimse<-function(x,tr=.2,na.rm=FALSE){
   if(na.rm)x<-x[!is.na(x)]
   trimse <- sqrt(var(winsorize(x,tr)))/((1-2*tr)*sqrt(length(x)))
   trimse
}

nboot <- 2000
alpha <- .95

sims <- matrix(sample(x, size = x*nboot, replace =T), nrow = nboot)
data <- sims - xbar

top <- apply(data, 1, mean)
bot <- apply(data, 1, trimse)

tval <- top/bot
tval <- sort(tval)

icrit <- round((1-alpha)*nboot)

c(Lower = mean(x)+tval[icrit]*trimse(x),
   Upper = mean(x)-tval[icrit]*trimse(x))
```

Why do they differ?

_Student's T gives a shorter confidence interval, however the probability coverage suggest less than 95% coverage._

### 6.15

Rats are subjected to a drug that might affect aggression. Suppose that for a random sample of rats, measures of aggression are found to be:

5, 12, 23, 24, 18, 9, 18, 11, 36, 15

Compute a 0.95 confidence interval for the mean assuming the scores are from a normal distribution.

```{r, echo = T}
x <- c(5, 12, 23, 24, 18, 9, 18, 11, 36, 15)

xbar <- mean(x); n <- length(x)

xbar + qt(conf(.95), df = n - 1) * sd(x)/sqrt(n)
```

### 6.16

Suppose $M = 34$ and the McKean-Schrader estimate of the standard error of M is $S_M = 3$.

Compute a .95 confidence interval for the population median.

```{r, echo = T}
34 + qnorm(conf(.95)) * 3
```

### 6.17

For the data in E14, the McKean-Schrader estimate of the standard error of _M_ is $S_M = 77.8$ and the sample median is 262.

Compute a .99 confidence interval for the population median.

```{r, echo = T}
x <- c(77,87,88,114,151,210,219,246,253,262,296,299,306,376,428,515, 666,1310,2611)
M <- median(x); n <- length(x)

M + qnorm(conf(.99)) * 77.8
```

### 6.18

If $n = 10$ and a confidence interval for the median is computed using $(X_k, X_{n-k})$ as described in 6.3.4, what is the probability coverage if $k=2$?

```{r, echo = T}
1 - 2*pbinom(1, size = 10, prob = .5)
```

### 6.19

From the previous, what is the probability coverage when  $n = 15$ and $k = 4$?

```{r, echo = T}
1 - 2*pbinom(3, size = 15, prob = .5)
```

### 6.20

For the data in E14, if you use (88, 515) as a confidence interval for the median, what is the probability this interval contains the population median?

```{r, echo = T}
x <- c(77, 87, 88, 114, 151, 210, 219, 246, 253, 262, 296, 299, 306, 376, 428, 515, 666, 1310, 2611)

k <- 3

1 - 2*pbinom(k, size = n - k - 1, prob = .5)
```

### 6.21

You observe the following successes and failures:

1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0

Compute a .95 confidence interval for $p$ using E6.15, as well as the Agresti-Coull method.

```{r, echo = T}
x <- c(1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
n <- length(x)

alpha <- 0.05

phat <- sum(x) / n

phat + qnorm(conf(1 - alpha)) * sqrt(phat*(1 - phat)/n)

c <- qnorm(1 - alpha/2)
nhat <- n + c^2
xhat <- sum(x) + c^2/2
phat <- xhat / nhat

c(Lower = phat - c * sqrt(phat*(1 - phat)/nhat),
  Upper = phat + c * sqrt(phat*(1 - phat)/nhat))
```

### 6.22

Given the following results for a sample from a binomial distribution, compute the squared standard error of $\hat{p}$.

a.) $n = 25, X = 5$

```{r, echo = T}
n <- 25; success <- 5
phat <- success/n

phat*(1 - phat)/n
```

b.) $n = 48, X = 12$

```{r, echo = T}
n <- 48; success <- 12
phat <- success/n

phat*(1 - phat)/n
```

c.) $n = 100, X = 80$

```{r, echo = T}
n <- 100; success <- 80
phat <- success/n

phat*(1 - phat)/n
```

d.) $n = 300, X = 160$

```{r, echo = T}
n <- 300; success <- 160
phat <- success/n

phat*(1 - phat)/n
```

### 6.23

Among 100 randomly sampled adults, 10 were found to be unemployed.

Compute a 0.95 confidence interval for the percentage of adults unemployed using E6.15.

Compare this result to the Agresti-Coull confidence interval.

### 6.24

A sample of 1,000 fishes was obtained from a lake. It was found that 290 were members of the bass family.

Compute a 0.95 confidence interval for the percenttage of bass in the lake using E6.15.

### 6.25

Among a random sample of 1,000 adults, 60 reported never having any legal problems. Use Equation 6.15 to compute the .95 confidence interval for the percentage of adults who have never had legal problems.

### 6.26

Among a sample of 600 items, only one was found to be defective. Explain why E6.15 might be unsatisfactory when computing a confidence interval for the probability that an item is defective.

### 6.27

In the previous exercise, compute a 0.90 confidence interval for the probability that an item is defective.

### 6.28

One-fourth of 300 persons in a large city stated that they are opposed to a certain political issue favored by the mayor.

Using E6.15, calculate a .99 confidence interval for the fraction of individuals opposed to this issue.

### 6.29

Consider a method for detecting a type of cancer that is highly accurate but very expensive and invasive, resulting in a great deal of discomfort for the patient. An alternative test to detect this type of cancer has been developed, and it is of interest to know the probability of a false-negative indication, meaning that the test fails to detect cancer when it is present. The test is conducted on 250 patients known to have cancer, and 5 tests fail to show its presence. Use equation 6.15 to determine a 0.95 confidence interval for the probability of a false-negative indication.

### 6.30

In the previous exercise, imagine that 0 false-negative indications were found.

Determine a .99 confidence interval for the probability of a false-negative indication.

### 6.31

### 6.32

### 6.33

### 6.34

### 6.35

### 6.36

### 6.37

### 6.38

### 6.39

### 6.40

### 6.41

### 6.42