---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Hypothesis Testing}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 7}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(here, quietly = T, warn.conflicts = F)
library(ggplot2, quietly = T, warn.conflicts = F)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(WRS, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

install.packages("WRS")
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 7

### 7.1

Given that $\bar{X} = 78, \sigma^2 =25, n = 10, \alpha = 0.05$ test $H_0 : \mu \ge 80$, assuming that observations are randomly sampled from a normal distribution.

```{r, echo = T}
xbar <- 78; sigma <- 5; n <- 10
mu <- 80; alpha <- 0.05

# mu > 80 = left tail test

crit <- qnorm(alpha)

Z <- sqrt(n)*(xbar - mu)/sigma

ifelse(Z > crit, "Fail to reject the null", "Reject the null")

```

### 7.2

Repeat the previous exercise, but test $H_0 : \mu = 80$

```{r, echo = T}
xbar <- 78; sigma <- 5; n <- 10
mu <- 80; alpha <- 0.05

# mu = 80, two-tailed test

crit <- qnorm(1 - alpha/2)

Z <- sqrt(n)*(xbar - mu)/sigma

ifelse(abs(Z) <= crit, "Fail to reject the null", "Reject the null")

```

### 7.3

For the data in Exercise 1, compute a 0.95 confidence interval and verify that this interval is consistent with y our decision about whether to reject the null hypothesis $H_0 : \mu = 80$.

```{r, echo = T}
xbar <- 78; sigma <- 5; n <- 10
mu <- 80; alpha <- 0.05

xbar + qnorm(c(Lower = alpha/2, Upper = 1 - (alpha/2))) * sigma/sqrt(n)
```

_80_ is in the confidence interval, so fail to reject.

### 7.4

For exercise 1, determine the p-value.

### 7.5

For exercise 2, determine the p-value.

### 7.6

Given that $\bar{X} = 120, \sigma = 5, n = 49, \alpha = 0.05$, test $H_0 : \mu \ge 130$, assuming that observations are randomly sampled from a normal distribution.

### 7.7

Repeat the previous exercise but test $H_0 : \mu = 130$.

### 7.8

For the previous exercise, compute a 0.95 confidence interval and compare the result with your decision about whether to reject $H_0$.

### 7.9

If $\bar{X} = 23$ and $\alpha = 0.025$, can you make a decision about whether to reject $H_0 : \mu \le 25$ without knowing $\sigma$?

### 7.10
### 7.11
### 7.12
### 7.13
### 7.14
### 7.15
### 7.16
### 7.17
### 7.18
### 7.19

### 7.20
### 7.21
### 7.22
### 7.23
### 7.24
### 7.25
### 7.26
### 7.27
### 7.28
### 7.29

### 7.30
### 7.31
### 7.32
### 7.33
### 7.34
### 7.35
### 7.36
### 7.37

