---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Interpretable Machine Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 16}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---


```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(kernlab, quietly = TRUE, warn.conflicts = FALSE)
library(mlbench, quietly = TRUE, warn.conflicts = FALSE)

library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(iml, quietly = TRUE, warn.conflicts = FALSE)
library(DALEX, quietly = TRUE, warn.conflicts = FALSE)
library(lime, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Interpretable Machine Learning

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

set.seed(123)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

```{r}
# ensure consistent categorical levels
blueprint <- recipe(Sale_Price ~., data = ames_train) %>%
  step_other(all_nominal(), threshold = 0.005)

# Create training / test h2o frames
train_h2o <- prep(blueprint, training = ames_train, retain = T) %>%
  juice() %>%
  as.h2o()

test_h2o <- prep(blueprint, training = ames_train) %>%
  bake(new_data = ames_test) %>%
  as.h2o()

Y <- "Sale_Price"
X <- setdiff(names(ames_train), Y)
```

### Stacking Overview

Stacking is the process of combining multiple "base" learners (RF, GBM, GLM, etc.) into a "Super Learner" that uses the individual learners to make the final prediction.

The Super Learner Algorithm

1.) Setup the Ensemble

  Specify a list of L base learners (with a specific set of model parameters)
  
  Specify a meta learning algorithm. This can be any one of the algorithms discussed in the previous chapters, but most often is some form of regularized regression.
  
2.) Train the ensemble

  Train each of the L base learners on the training set.
  
  Perform k-fold CV on each of the base learners and collect the cross-validated predictions from each (the same k-fold must be used for each base learner). These predicted values represent $p_1, ..., p_l$.
  
  The N cross-validated predicted values from each of the L algorithms can be combined to form a new N X L feature matrix (Z)
  
  Train the meta learning algorithm on level-one data (y = f(Z)). The "ensemble model" consista of the L base learning models and the meta learning model, which can then be used to generate predictions on new data.
  
3.) Predict on new data.

  To generate ensemble predictions, first generate predictions from the base learners.
  
  Feed those predictions into the meta learner to generate the ensemble prediction.

### Stacking Existing Models

Restrictions:

1.) All models must be trained on the same training set

2.) All models must be trained with the same number of CV folds.

3.) All models must use the same fold assignment to ensure the same observations are used.

4.) The cross-validated predictions from all of the models must be preserved by setting keep_cross_validation_predictions = T.


```{r}
# Train & cross-validate a GLM model
best_glm <- h2o.glm(
  x = X, y = Y, training_frame = train_h2o, alpha = 0.1,
  remove_collinear_columns = TRUE, nfolds = 10, fold_assignment = "Modulo",
  keep_cross_validation_predictions = TRUE, seed = 123
)

# Train & cross-validate a RF model
best_rf <- h2o.randomForest(
  x = X, y = Y, training_frame = train_h2o, ntrees = 1000, mtries = 20,
  max_depth = 30, min_rows = 1, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate a GBM model
best_gbm <- h2o.gbm(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.01,
  max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,es
  
  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate an XGBoost model
best_xgb <- h2o.xgboost(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
  nfolds = 10, fold_assignment = "Modulo", 
  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
  stopping_metric = "RMSE", stopping_tolerance = 0
)
```

Train the ensemble

```{r}
# Train a stacked tree ensemble
ensemble_tree <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble",
  base_models = list(best_glm, best_rf, best_gbm, best_xgb),
  metalearner_algorithm = "drf"
)
```

```{r}
get_rmse <- function(model) {
  results <- h2o.performance(model, newdata = test_h2o)
  results@metrics$RMSE
}

list(best_glm, best_rf, best_glm, best_xgb) %>%
  purrr::map_dbl(get_rmse)
```

Stacked Results

```{r}
h2o.performance(ensemble_tree, newdata = test_h2o)@metrics$RMSE
```

```{r}
data.frame(
  GLM_pred = as.vector(h2o.getFrame(best_glm@model$cross_validation_holdout_predictions_frame_id$name)),
  RF_pred = as.vector(h2o.getFrame(best_rf@model$cross_validation_holdout_predictions_frame_id$name)),
  GBM_pred = as.vector(h2o.getFrame(best_gbm@model$cross_validation_holdout_predictions_frame_id$name)),
  XGB_pred = as.vector(h2o.getFrame(best_xgb@model$cross_validation_holdout_predictions_frame_id$name))
) %>% cor()
```

### Stacking a Search Grid

```{r}
# Define GBM hyperparameter grid
hyper_grid <- list(
  max_depth = c(1, 3, 5),
  min_rows = c(1, 5, 10),
  learn_rate = c(0.01, 0.05, 0.1),
  learn_rate_annealing = c(0.99, 1),
  sample_rate = c(0.5, 0.75, 1),
  col_sample_rate = c(0.8, 0.9, 1)
)

# Define random grid search criteria
search_criteria <- list(
  strategy = "RandomDiscrete",
  max_models = 25
)

# Build random grid search 
random_grid <- h2o.grid(
  algorithm = "gbm", grid_id = "gbm_grid", x = X, y = Y,
  training_frame = train_h2o, hyper_params = hyper_grid,
  search_criteria = search_criteria, ntrees = 5000, stopping_metric = "RMSE",
  stopping_rounds = 10, stopping_tolerance = 0, nfolds = 10, 
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123
)
```

```{r}
# Sort results by RMSE
h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "rmse"
)
```

```{r}
# Grab the model_id for the top model, chosen by validation error
best_model_id <- random_grid@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)
h2o.performance(best_model, newdata = test_h2o)
```

```{r}
# Train a stacked ensemble using the GBM grid
ensemble <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "ensemble_gbm_grid",
  base_models = random_grid@model_ids, metalearner_algorithm = "gbm"
)

# Eval ensemble performance on a test set
h2o.performance(ensemble, newdata = test_h2o)
```

### Auto ML

```{r}
# Use AutoML to find a list of candidate models (i.e., leaderboard)
auto_ml <- h2o.automl(
  x = X, y = Y, training_frame = train_h2o, nfolds = 5, 
  max_runtime_secs = 60 * 120, max_models = 50,
  keep_cross_validation_predictions = TRUE, sort_metric = "RMSE", seed = 123,
  stopping_rounds = 50, stopping_metric = "RMSE", stopping_tolerance = 0
)
```

```{r}
# Assess the leader board; the following truncates the results to show the top 
# 25 models. You can get the top model with auto_ml@leader
auto_ml@leaderboard %>% 
  as.data.frame() %>%
  dplyr::select(model_id, rmse) %>%
  dplyr::slice(1:25)
```

#### Clean-up

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```

