---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
   \rfoot{\color{headergrey}Chapter 4}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
library(gtools, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### 4.1

Consider the population {1, 2, 5, 6, 10, 12}.

Find (and plot) the sampling distribution of medians for samples of size 3 without replacement.

```{r, echo = T, fig.height=4.5, fig.width=8}

p <- c(1, 2, 5, 6, 10, 12)
c <- combinations(v = p, n = 6, r = 3)
t <- apply(c, 1, median)

ggplot(data.table(value = t), aes(value, fill = ..count..)) +
   geom_histogram(bins = 30) +
   labs(title = "Median Sampling Distribution of p")

```

Compare the median of the population to the mean of the medians.

Median of p = __`r median(p)`__. Mean of Medians of p = __`r mean(t)`__

\newpage

### 4.2

Consider the population {3, 6, 7, 9, 11, 14}.

For samples of size 3 without replacement, find (and plot) the sampling distribution for the minimum.

```{r, echo = T, fig.height=4.5, fig.width=8}

p <- c(3, 6, 7, 9, 11, 14)
c <- combinations(v = p, n = 6, r = 3)
t <- apply(c, 1, min)

ggplot(data.table(value = t), aes(value, fill = ..count..)) +
   geom_histogram(bins = 30) +
   labs(title = "Minimum Sampling Distribution of p")

```

What is the mean of the sampling distribution? __`r mean(t)`__

The statistic is an estimate of some parameter - what is the value of that parameter?

This is an estimation of the minimum, which is: __`r min(p)`__

\newpage

### 4.3

Let _A_ denote the population {1, 3, 4, 5} and _B_ the population {5, 7, 9}.

```{r, echo = T}

A <- c(1, 3, 4, 5)
B <- c(5, 7, 9)

```

Let _X_ be a random value from _A_, and _Y_ and random value from _B_.

a.) Find the sampling distribution of _X + Y_.

```{r, echo = T}

result = numeric(12)
index <- 1
for(j in 1:length(A))
{
   for(k in 1:length(B))
   {
      result[index] <- A[j] + B[k]
      index <- index + 1
   }
}

sort(result)

```

b.) In this example, does the sampling distribution depend on whether you sample with or without replacement?

_No._

Why or why not?

_Because 5 in is both sets._

c.) Compute the mean of the values for each of _A_ and _B_ and the values in the sampling distribution of _X + Y_.

Mean of _A_: __`r mean(A)`__. Mean of _B_: __`r mean(B)`__.

Mean of _A + B_: __`r mean(result)`__

How are the means related?

mean(A) + mean(B) = mean(A + B).

d.) Suppose you draw a random value from _A_ and a random value from _B_.

```{r, echo = T}

prob <- sum(result >= 13) / length(result)

```

What is the probability that the sum is 13 or larger? __`r prob*100`%__

\newpage

### 4.4

Consider the population {3, 5, 6, 6, 8, 11, 13, 15, 19, 20}.

a.) Compute the mean and standard deviation and create a dot plot of its distribution.

```{r, echo = T, fig.height=3.5, fig.width=8}

p <- c(3, 5, 6, 6, 8, 11, 13, 15, 19, 20)

mu <- mean(p)
sigma <- sd(p)

ggplot(data.table(value = p)) +
   geom_dotplot(aes(value, fill = ..count..), binwidth = 1) +
   labs(title = "Population Dot Plot")

```

$\mu = `r mu`$, $\sigma = `r sigma`$

b.) Simulate the sampling distribution of $\bar{X}$ by taking random samples of size 4 and plot your results.

```{r, echo = T, fig.height=3.5, fig.width=8}

N <- 10e2
results <- numeric(N)

for( i in 1:N)
{
   index <- sample(length(p), size = 4, replace = F)
   results[i] <- mean( p[index] )
}

ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Sample Means")

xbar <- mean(results)
se <- sd(results) / sqrt(N)

```

Compute the mean and standard error, and compare to the population mean and standard deviation.

mean: `r xbar`, standard error: `r se`

c.) Use the simulation to find $P(\bar{X} < 11)$.

```{r, echo = T}

prob <- mean(results < 11)

```

$P(\bar{X} < 11) = `r prob*100`$%

\newpage

### 4.5

Consider two populations A = {3, 5, 7, 9, 10, 16}, B = {8, 10, 11, 15, 18, 25, 28}.

```{r, echo = T}
A <- c(3, 5, 7, 9, 10, 16)
B <- c(8, 10, 11, 15, 18, 25, 28)
```

a.) Using R, draw random samples (without replacement) of size 3 from each population, and simulate the sampling distribution of the sum of their maximums.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp.a <- sample(A, 3, replace = F)
   samp.b <- sample(B, 3, replace = F)
   
   results[i] <- max(samp.a) + max(samp.b) 
}

ggplot(data.table(value = results)[, index := .I]) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Sampling Distribution: max(A) + max(B)")

``` 

b.) Use your simulation to estimate the probability that the sum of the maximums is less than 20.

```{r, echo = T}
prob <- mean(results < 20)
```

Probability: `r prob*100`%

c.) Draw random samples of size 3 from each population, and find the maximum of the union of these two sets.

Simulate the sampling distribution of the maximums of this union.

```{r, echo = T, fig.height=3.5, fig.width=8}
results <- numeric(N)

for(i in 1:N)
{
   samp.a <- sample(A, 3, replace = F)
   samp.b <- sample(B, 3, replace = F)
   
   results[i] <- max(union(samp.a, samp.b))
}

ggplot(data.table(value = results)[, index := .I]) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Sampling Distribution: max(union(A,B))")
```

d.) Use simulation to find the probability that the maximum of the union is less than 20.

```{r, echo = T}
prob <- mean(results < 20)
```

Probability: `r prob*100`%

\newpage

### 4.6

The data set _Recidivism_ contains the poopulation of all Iowa offenders convicted of either a felony or misdemeanor who were released in 2010 (case study in Section 1.4).

```{r, echo = T}
Recidivism <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
                               header = T))
```

Of these, 31.6% recidivated and were sent back to prision.

Simulate the sampling distribution of $\hat{p}$, the sample proportion of offeneders who recidivated, for random samples of size 25.

```{r, echo = T}

mean(Recidivism$Recid == "Yes")

N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Recidivism$Recid, 25)
   results[i] <- mean(samp == "Yes")
}

```

a.) Create a histogram and describe the simulated sampling distribution of $\hat{p}$.

```{r, echo = T, fig.height=3.5, fig.width=8}
ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Recidivism Sampling Distribution")
```

Estimate the mean and standard error.

```{r, echo = T}
mu <- mean(results)
se <- sd(results) / sqrt(25)
```

$\mu = `r mu`$, $\sigma = `r se`$

b.) Compare your estimate of the standard error with the theoretical standard error (_Corollary 4.3.2_).

```{r, echo = T}
tse <- mu * ( 1 - mu ) / sqrt(25)
```

Theoretical: `r tse`

c.) Repeat the above using samples of size 250, and compare with the $n = 25$ case.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Recidivism$Recid, 250)
   results[i] <- mean(samp == "Yes")
}

ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Recidivism Sampling Distribution")

mu <- mean(results)
se <- sd(results) / sqrt(250)

```

$\mu = `r mu`$, $\sigma = `r se`$

### 4.7

The data set _FlightDelays_ contains the population of all flight departures by United Airlines and American Airlines out of LGA during May and June 2009 (case study in Section 1.1).

```{r, echo = T}
Flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
                               header = T))
```

a.) Create a histogram of _Delay_ and describe the distribution.

```{r, echo = T, fig.height=3.5, fig.width=8}

ggplot(Flights, aes(Delay)) +
   geom_histogram(aes(fill = ..count..), bins = 30) +
   scale_y_continuous(labels = comma) +
   labs(title = "Flight Delays")

```

Compute the mean and standard deviation.

```{r, echo = T}
mu <- mean(Flights$Delay)
sigma <- sd(Flights$Delay)
```

$\mu = `r mu`$, $\sigma = `r sigma`$

b.) Simulate the sampling distribution of $\bar{x}$, the sample mean of the length of the flight delays (_Delay_), for sample size 25.

```{r, echo = T}

N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Flights$Delay, 25, replace = F)
   results[i] <- mean(samp)
}

```

Create a histogram and describe the simulated sampling distribution of $\bar{x}$.

```{r, echo = T, fig.height=3.5, fig.width=8}
ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Flight Delay Sampling Distribution")
```

Estimate the mean and standard error.

```{r, echo = T}
mu <- mean(results)
se <- sd(results) / sqrt(25)
```

$\mu = `r mu`$, $\Sigma = `r se`$

c.) Compare your estimate of the standard error with the theoretical standard error (_Corollary A.4.1_).

```{r, echo = T}
tse <- var(results) / 25
```

Theoretical: `r tse`

d.) Repeat with sample size 250.

```{r, echo = T, fig.height=3.5, fig.width=8}

N <- 10e2
results <- numeric(N)

for(i in 1:N)
{
   samp <- sample(Flights$Delay, 250, replace = F)
   results[i] <- mean(samp)
}

ggplot(data.table(value = results)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
   labs(title = "Flight Delay Sampling Distribution")

mu <- mean(results)
se <- sd(results) / sqrt(250)
tse <- var(results) / 250
```

$\mu = `r mu`$, $\Sigma = `r se`$

Theoretical: `r tse`

### 4.8

Let $X_1, X_2, \ldots, X_{25}$ be a random sample from some distribution and $W = T(X_1, X_2, \ldots, X_n)$ be a statistic.

Suppose the _sampling distribution_ of Y has pdf $f(y) = (3/8)y^2 \ for \ 0 \le y \le 2$.

Find $P(0 \le Y \le \frac{1}{5})$

