---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Cherry Blossom Race Times}
   \rfoot{\color{headergrey}Chapter 2}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(codetools, quietly = TRUE, warn.conflicts = FALSE)
library(lattice, quietly = TRUE, warn.conflicts = FALSE)
library(fields, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)

library(rbenchmark, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- file.path(here::here(), "Case Studies", "datasets")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Cherry Blossom Ten Mile Run & Walk

## Examining the Impact of Age on Physical Performance

### The Raw Data

data is pre-processed from the web in \02_data.r

```{r}

load_results <- function(class, year) {
   
   folder <- ifelse(class == "men", "men_txt", "women_txt")
   file <- file.path(data.dir, folder, paste(year, ".txt", sep = ""))
   
   return(readLines(file))
}

```

### Race Results Preprocessing

```{r}

els <- load_results("men", 2006)

eqIndex <- grep("^===", els)
eqIndex

spacerRow <- els[eqIndex]
headerRow <- els[eqIndex - 1]

body <- els[ -(1:eqIndex) ]

headerRow <- tolower(headerRow)

timeStart <- regexpr("net", headerRow)
time <- substr(body, start = timeStart, stop = timeStart + 1)

ageStart <- regexpr("ag", headerRow)

age <- substr(body, start = ageStart, stop = ageStart + 1)
head(age)

summary(as.numeric(age))

blankLocs <- gregexpr(" ", spacerRow)
blankLocs

searchLocs <- c(0, blankLocs[[1]])

Values <- mapply(substr, list(body),
                 start = searchLocs[ -length(searchLocs)] + 1,
                 stop = searchLocs[ -1 ] - 1)

findColLocs <- function(spacerRow) {
   spaceLocs <- gregexpr(" ", spacerRow)[[1]]
   rowLength <- nchar(spacerRow)
   
   if(substring(spacerRow, rowLength, rowLength + 1) != " ")
      return(c(0, spaceLocs, rowLength + 1))
   else
      return(c(0, spaceLocs))
}

findColLocs(spacerRow)
name <- "home"
colnames <- c("name", "home", "ag", "gun", "net", "time")

colIndex <- which(colnames == name)

startPos <- regexpr(name, headerRow)[[1]]

#
# can we modify select cols to short circuit on matching w/o spaces?
#######


selectCols =
   function(colNames, headerRow, searchLocs)
   {
      sapply(colNames,
             function(name, headerRow, searchLocs)
            {
               startPos <- regexpr(name, headerRow)[[1]]
               
               if(startPos == -1)
                  return( c(NA, NA) )
               
               index <- sum(startPos >= searchLocs)
               
               c(searchLocs[index] + 1, searchLocs[index + 1])
            },
      headerRow = headerRow, searchLocs = searchLocs)
   }

searchLocs <- findColLocs(spacerRow)
loc <- selectCols("home", headerRow, searchLocs)
vars <- mapply(substr, list(body), start = loc[1,], stop = loc[2, ])

summary(as.numeric(vars))

shortColNames <- c("name", "home", "ag", "gun", "net", "time")

locCols <- selectCols(shortColNames, headerRow, searchLocs)
Values <- mapply(substr, list(body), start = locCols[1, ],
                 stop = locCols[2, ])

class(Values)

colnames(Values) = shortColNames

head(Values)
tail(Values)[, 1:3]
```

```{r}

extractVariables =
   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
   {
      # Find the index of the row with =s
      eqIndex <- grep("^===", file)
      spacerRow <- file[eqIndex]
      headerRow <- tolower(file[ eqIndex - 1 ])
      body <- file[ -(1 : eqIndex) ]
      
      blank <- grep("^[[:blank:]]*$", body)
      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
      ignore <- union(blank, footnote)
      
      if(length(ignore))
         body <- body[-ignore]
      
      # Obtain the starting and ending positions of variables
      searchLocs <- findColLocs(spacerRow)
      locCols <- selectCols(varNames, headerRow, searchLocs)
      
      Values <- mapply(substr, list(body), start = locCols[1, ],
                       stop = locCols[2, ])
      colnames(Values) <- varNames
      
      invisible(Values)
   }

load_data <- function(class) {
   folder <- ifelse(class == "men", "men_txt", "women_txt")
   filenames <- file.path(data.dir, paste(folder, "/", 1999:2012, ".txt", sep = ""))
   files <- lapply(filenames, readLines)
   names(files) <- 1999:2012
   
   mat <- lapply(files, extractVariables)
   
   return(mat)
}

```

### Data Cleaning and Reformatting Variables

```{r}
menResMat <- load_data("men")

length(menResMat)
sapply(menResMat, nrow)

womenResMat <- load_data("women")

length(womenResMat)
sapply(womenResMat, nrow)
```

Age Validation

```{r}
age <- as.numeric(menResMat[['2012']][, 'ag'])

tail(age)

age <- sapply(menResMat,
              function(x) as.numeric(x[, "ag"]))

age_values <- plyr::ldply(age, data.frame)
colnames(age_values) <- c("Year", "Age")

ggplot(age_values, aes(Age)) +
   geom_histogram(aes(fill = ..count..), bins = 30)

ggplot(age_values, aes(Year, Age)) +
   geom_boxplot(aes(fill = Year)) +
   labs(title = "Age By Race Year")
```

```{r}
sapply(age, function(x) sum(is.na(x)))

file2001 <- load_results("men", 2001)
age2001 <- as.numeric(extractVariables(file2001)[, "ag"])

badAgeIndex <- which(is.na(age2001)) + 5

file2001[badAgeIndex]

badAgeIndex

blanks <- grep("^[[:blank:]]*$", file2001)
blanks

which(age2001 < 5)

file2001[which(age2001 < 5)]


file <- load_results("men", 2001)
data <- extractVariables(file)
age <- as.numeric(data[, "ag"])

sum(age < 5)

which(age == 0)

data[which(age == 0)]

```

### Time Validation

```{r}
file2002 <- load_results("men", 2012)

charTime <- menResMat[["2012"]][, "time"]

head(charTime)
tail(charTime)

timePieces <- strsplit(charTime, ":")

timePieces[[1]]
tail(timePieces, 1)

timePieces <- sapply(timePieces, as.numeric)

runTime <- sapply(timePieces,
                  function(x) {
                     if(length(x) == 2) x[1] + x[2]/60
                     else 60 * x[1] + x[2] + x[3]/60
                  })

summary(runTime)

convertTime <- function( charTime ) {

   timePieces <- strsplit(charTime, ":")

   timePieces <- sapply(timePieces, as.numeric)

   runTime <- sapply(timePieces,
                  function(x) {
                     if(length(x) == 2) x[1] + x[2]/60
                     else 60 * x[1] + x[2] + x[3]/60
                  })
   
}

```

### Aggregate cleaning into 

```{r}

createDF =
   function(Res, year, sex) {
      useTime <- if(!is.na(Res[1, 'net']) )
                     Res[, 'net']
                  else if( !is.na(Res[1, 'gun']) )
                     Res[ , 'gun']
                  else
                     Res[, 'time']
      
      useTime <- gsub("[#\\*[:blank:]]", "", useTime)
      
      Res <- Res[ useTime != "", ]
      
      runTime <- convertTime(useTime[ useTime != "" ])
      
      N <- nrow(Res)
      
      Results <- data.frame( year = rep(year, N),
                             sex = rep(sex, N),
                             name = Res[, 'name'],
                             home = Res[, 'home'],
                             age = as.numeric(Res[, 'ag']),
                             runTime = runTime,
                             stringsAsFactors = F)
      
      invisible(Results)
   }

years <- 1999:2012
menDF <- mapply(createDF, menResMat, year = years,
                sex = rep("M", 14), SIMPLIFY = F)

warnings()[ c(1:2, 49:50)]

sapply(menDF, function(x) sum(is.na(x$runTime)))

file2006 <- load_results("men", 2006)
parsed2006 <- extractVariables(file2006)
time2006 <- parsed2006[, "net"]

womenDF <- mapply(createDF, womenResMat, year = years,
                  sex = rep("F", 14), SIMPLIFY = F)

sapply(womenDF, function(x) sum(is.na(x$runTime)))
```

### Consolidate Results

```{r}
cbMen <- do.call(rbind, menDF)
save(cbMen, file = file.path(data.dir, "cbMen.rda"))

cbWomen <- do.call(rbind, womenDF)

save(cbWomen, file = file.path(data.dir, "cbWomen.rda"))
```

### Exploratory Data Analysis

```{r}

ggplot(cbMen, aes(age, runTime)) +
   geom_point(aes(col = age)) +
   labs(main = "Run Times by Age", xlab = "Age (years)", xlab = "Run Time (minutes")

```

```{r}
display.brewer.all()
```

```{r}
Purple8 <- brewer.pal(9, "Purples")[8]

Purple8A <- paste(Purple8, "14", sep = "")

ggplot(cbMen, aes(age, runTime)) +
   geom_jitter(col = Purple8A) +
   geom_smooth() +
   xlim(10, 75) +
   ylim(45, 150)

```

```{r}
smoothScatter(y = cbMen$runTime, x = cbMen$age,
              ylim = c(40, 165), xlim = c(15, 85),
              xlab = "Age (years)", ylab = "Run Time (minutes)")
```

```{r}
cbMenSub <- cbMen[cbMen$runTime > 30 &
                     !is.na(cbMen$age) & cbMen$age > 15,]

ageCat <- cut(cbMenSub$age, breaks = c(seq(15, 75, 10), 90))

ggplot(data.table(age = ageCat, time = cbMenSub$runTime), aes(ageCat, time)) +
   geom_boxplot(aes(fill = age)) +
   labs("Run Times by Age Category")
```

```{r}
lm_age <- lm(runTime ~ age, data = cbMenSub)

summary(lm_age)
```

```{r}
smoothScatter(x = cbMenSub$age, y = lm_age$residuals,
              xlab = "Age (years)", ylab = "Residuals") 
abline(h = 0, col = "purple", lwd = 3)

resid.lo <- loess(resids ~ age,
                  data = data.frame(resids = residuals(lm_age),
                                    age = cbMenSub$age))

age20to80 <- 20:80

resid.lo.pr <- predict(resid.lo, newdata = data.frame(age = age20to80))

lines(x = age20to80, y = resid.lo.pr,
      col = "green", lwd = 3, lty = 2)
```

```{r}
menRes.lo <- loess(runTime ~ age, cbMenSub)

summary(menRes.lo)

menRes.lo.pr <- predict(menRes.lo, data.frame(age = age20to80))

```

```{r}
over50 <- pmax(0, cbMenSub$age - 50)

lmOver50 <- lm(runTime ~ age + over50, data = cbMenSub)

summary(lmOver50)

decades <- seq(30, 60, by = 10)
overAge <- lapply(decades,
                  function(x) pmax(0, (cbMenSub$age - x)))
names(overAge) <- paste("over", decades, sep = "")
overAge <- as.data.frame(overAge)

tail(overAge)

lmPiecewise <- lm(runTime ~., 
                  data = cbind(cbMenSub[, c("runTime", "age")],
                               overAge))

summary(lmPiecewise)

overAge20 <- lapply(decades, function(x) pmax(0, (age20to80 - x)))
names(overAge20) <- paste("over", decades, sep = "")
overAgeDF <- cbind(age = data.frame(age = age20to80), overAge20)

tail(overAgeDF)
```

```{r}
predPiecewise <- predict(lmPiecewise, overAgeDF)

plot(predPiecewise ~ age20to80,
     type = "l", col = "purple", lwd = 3,
     xlab = "Age (years)", ylab = "Run Time Prediction")

lines(x = age20to80, y = menRes.lo.pr,
      col = "green", lty = 2, lwd = 3)

legend("topleft", col = c("purple", "green"),
       lty = c(1, 2), lwd = 3,
       legend = c("Piecewise Linear", "Loess Curve"), bty = "n")
```

### Cross-Sectional Data and Covariates

```{r}
age1999 <- cbMenSub[ cbMenSub$year == 1999, "age"]
age2012 <- cbMenSub[ cbMenSub$year == 2012, "age"]
```

```{r}
summary(age1999)

summary(age2012)
```

```{r}
plot(density(age1999, na.rm = T),
     ylim = c(0, 0.05), col = "purple",
     lwd = 3, xlab = "Age (years)", main = "")
lines(density(age2012, na.rm = T),
      lwd = 3, lty = 2, col = "green")
legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
       legend = c("1999", "2012", bty = "n"))
```

```{r}
by_year <- data.table(cbMenSub)[, .(age, mean = mean(age)), by = list(year)]

ggplot(by_year[year %in% c("1999", "2012")], aes(age, group = year)) +
   geom_density(aes(fill = year), alpha = .6) +
   geom_vline(aes(xintercept = mean, col = year, group = year)) +
   labs(title = "Runners by Year")
```

```{r}
ggplot(by_year, aes(year, age)) +
   geom_point(aes(col = age)) +
   geom_quantile() +
   labs(title = "Runners Age Over Time")
```

```{r}
mR.lo99 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 1999,])
mR.lo.pr99 <- predict(mR.lo99, data.frame(age = age20to80))

summary(mR.lo.pr99)

mR.lo12 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 2012, ])
mR.lo.pr12 <- predict(mR.lo12, data.frame(age = age20to80))

summary(mR.lo12)

plot(mR.lo.pr99 ~ age20to80,
     type = "l", col = "purple", lwd = 3,
     xlab = "Age (years)", ylab = "Fitted Run Time (minutes)")

lines(x = age20to80, y = mR.lo.pr12,
      col = "green", lty = 2, lwd = 3)

legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
       legend =  c("1999", "2012"), bty = "n")
```

```{r}

years <- 1999:2012
results <- list(length(years))

y <- 1999

for( i in 1:length(years) )
{
   y = years[i]
   data <- cbMenSub[ which(cbMenSub$year == y), ]
   
   model <- loess(runTime ~ age, data)
   
   pred <- predict(model, newdata = age20to80)
   
   results[[i]] <- data.table(Year = rep(y, length(pred)), Age = age20to80, Pred = pred)
}

race_years_data <- do.call("rbind", results)

ggplot(race_years_data, aes(Age, Pred, group = Year)) +
   geom_line(aes(col = Year)) +
   labs(title = "Predicted Performance by Year", ylab = "Predicted Run Time", xlab = "Runner Age")

```

### Constructing a Record for an Individual Runner across Years

```{r}
trimBlanks <- function(charVector) {
   nameClean <- gsub("^[[:blank:]]+", "", charVector)
   nameClean <- gsub("[[:blank:]]+$", "", nameClean)
   nameClean <- gsub("[[:blank:]]+", " ", nameClean)
}

nameClean <- trimBlanks(cbMenSub$name)

length(nameClean)

length(unique(nameClean))

table(table(nameClean))

```


```{r}
head(sort(table(nameClean), decreasing = T), 1)

mSmith <- cbMenSub[nameClean == "Michael Smith", ]

head(unique(mSmith$home))

nameClean <- tolower(nameClean)

head(sort(table(nameClean), decreasing = T), 1)

nameClean <- gsub("[,.]", "", nameClean)
```

```{r}
tabNameYr <- table(cbMenSub$year, nameClean)

max(tabNameYr)

class(tabNameYr)

mode(tabNameYr)

names(attributes(tabNameYr))

dim(tabNameYr)

head(colnames(tabNameYr), 3)

which(tabNameYr == max(tabNameYr), arr.ind = T)

indMax <- which(tabNameYr == max(tabNameYr), arr.ind = T)
colnames(tabNameYr)[indMax[2]]

cbMenSub$nameClean <- nameClean
```

```{r}
cbMenSub$yob <- cbMenSub$year - cbMenSub$age

homeClean <- trimBlanks(cbMenSub$home)
homeClean <- tolower(homeClean)

cbMenSub$homeClean <- homeClean
```

```{r}
vars <- c("year", "homeClean", "nameClean", "yob", "runTime")
mb <- which(nameClean == "michael brown")
birthOrder <- order(cbMenSub$yob[mb])
cbMenSub[mb[birthOrder], vars]
```

```{r}
cbMenSub$ID = paste(nameClean, cbMenSub$yob, sep = "_")

races <- tapply(cbMenSub$year, cbMenSub$ID, length)

races8 <- names(races)[which(races >= 8)]

men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]

orderByRunner <- order(men8$ID, men8$year)

men8 <- men8[orderByRunner, ]
```

```{r}
men8L <- split(men8, men8$ID)
names(men8L)
```

```{r}
length(unique(men8$ID))
length(men8L)
```

```{r}
gapTime <- tapply(men8$runTime, men8$ID,
                  function(t) any(abs(diff(t)) > 20))

gapTime <- sapply(men8L, function(df)
   any(abs(diff(df$runTime)) > 20))

sum(gapTime)
```

```{r}
lapply(men8L[ gapTime ][1:2], function(df) df[, vars])
```

```{r}
homeLen <- nchar(cbMenSub$homeClean)

cbMenSub$state <- substr(cbMenSub$homeClean,
                         start = homeLen - 1, stop = homeLen)

cbMenSub$state[cbMenSub$year == 2006] = NA


cbMenSub$ID = paste(cbMenSub$nameClean, cbMenSub$yob,
                    cbMenSub$state, sep = "_")
```

```{r}
numRaces <- tapply(cbMenSub$year, cbMenSub$ID, length)
races8 <- names(numRaces)[which(numRaces >= 8)]
men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]

orderByRunner <- order(men8$ID, men8$year)
men8 <- men8[orderByRunner, ]

men8L <- split(men8, men8$ID)
names(men8L) <- races8

length(men8L)
```

### Modeling the Change in Running Time for Individuals


```{r}
groups <- 1 + (1:length(men8L) %% 9)

plot(x = 40, y = 60, type = "n",
     xlim = c(20, 80), ylim = c(40, 160),
     xlab = "Age (years)", ylab = "Run Time (minutes)")


addRunners <- function(listRunners, colors, numLty) {
   numRunners <- length(listRunners)
   
   colIndx <- 1 + (1:numRunners) %% length(colors)
   ltys <- rep(1:numLty, each = length(colors), length = numRunners)
   
   mapply(function(df, i) {
      lines(df$runTime ~ df$age,
            col = colors[colIndex[i]], lwd = 2, lty = ltys[i])
   }, listRunners, i = 1:numRunners)
}

colors <- c("#e41alc", "#377eb8", "#4daf4a", "#984ea3",
            "#ff7f00", "#a65628")

par(mfrow = c(3, 3), mar = c(2, 2, 1, 1))
invisible(
   sapply(1:9, function(grpId) {
      plot(x = 0, y = 0, type = "n",
           xlim = c(20, 80), ylim = c(50, 130),
           xlab = "Age (years)", ylab = "Run Time (minutes)")
      
      addRunners(men8L[ groups == grpId ], colors, numLty = 6)
   })
)
```

```{r}
fitOne <- function(oneRunner, addLine = F, col = "grey") {
   lmOne <- lm(runTime ~ age, data = oneRunner)
   
   if ( addLine )
      lines(x = oneRunner$age, y = predict(lmOne),
            col = col, lwd = 2, lty = 2)
   
   ind <- floor( (nrow(oneRunner) + 1) / 2)
   res <- c(coefficients(lmOne)[2], oneRunner$age[ind],
            predict(lmOne)[ ind ])
   
   names(res) <- c("ageCoeff", "medAge", "predRunTime")
   
   return(res)
}
```

```{r}
par(mfrow = c(1, 1))
plot( x = 0, y = 0, type = "n",
      xlim = c(20, 80), ylim = c(50, 130),
      xlab = "Age (years)", ylab = "Run Time (minutes)")

addRunners( men8L[ groups == 9 ], colors, numLty = 6)
lapply(men8L[groups == 9], fitOne, addLine = T, col = "black")
```

```{r}
men8LongFit <- lapply(men8L, fitOne)

coeffs <- sapply(men8LongFit, "[", "ageCoeff" )
ages <- sapply(men8LongFit, "[", "medAge")

longCoeffs <- lm(coeffs ~ ages)

summary(longCoeffs)
```

### Further Analysis

### 1.)

Write a function that uses _read.fwf()_ to read the 28 text tables in MenTxt/ and WomenTxt/ into R. These are called 1999.text, 2000txt, etc. and are described in greater detail in 2.2. Examine the tables in a plain text editor to determine the start and end position of each column of interest (name, hometown, age and gun and net time).

Use statistics to explore the results and confirm that you have extracted the information from the correct positions in the text.

```{r}
data_paths <- sapply(c("men_txt", "women_txt"), function(p) {
   path <- file.path(data.dir, p)

   sapply(path, function( f ) { 
      file.path(path, list.files(path) )
   })
})

consolidated_files <- c(data_paths[, 1], data_paths[, 2])


tp <- consolidated_files[1]

files <- lapply(consolidated_files, function(f) read.fwf(f, widths = 120))

length(files)
```


### 2.)

Revise the extractVariables function (see section 2.2) to remove the rows in _menTables_ that are blank. In addition, eliminatee the rows that begin with a '*' or a '#'. You may find the following regular expression helpful for locating blank rows in a table.

_grep("^[[:blank:]]*$", body)_

The pattern uses several meta characters. The ^ is an anchor for the start of the string, the $ anchors to the end of the string, the [[:blank:]] denotes the equivalence class of any space or tab character, and * indicates that the blank character can appear 0 or more times. All together the pattern ^[[:blank:]]*\$ matches a string that contains any number of blanks from start to end.

```{r}

extractVariables =
   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
   {
      # Find the index of the row with =s
      eqIndex <- grep("^===", file)
      spacerRow <- file[eqIndex]
      headerRow <- tolower(file[ eqIndex - 1 ])
      body <- file[ -(1 : eqIndex) ]
      
      blank <- grep("^[[:blank:]]*$", body)
      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
      ignore <- union(blank, footnote)
      
      if(length(ignore))
         body <- body[-ignore]
      
      # Obtain the starting and ending positions of variables
      searchLocs <- findColLocs(spacerRow)
      locCols <- selectCols(varNames, headerRow, searchLocs)
      
      Values <- mapply(substr, list(body), start = locCols[1, ],
                       stop = locCols[2, ])
      colnames(Values) <- varNames
      
      invisible(Values)
   }

```

### 3.)

Find the record where the time is only 1.5. What happened? Determine how to handle the problem and which function needs to be modified: _extractResTable()_, _extractVariables()_, or _cleanUp()_. In your modification, include code to provide a warning message about the rows that are being dropped for having a time that is too small.

```{r}

```

### 4.)

Examine the head and tail of the 2006 men's file. Look at both the character matrix in the list called _menResMat_ and the character vector in the list called _menFiles_ (see Sec 2.2). (Recall that the desired character matrix in _menResMat_ and the character vector in _menFiels_ both correspond to the element named "2006"). What is wrong with the hometown? Examine the header closely to figure out how this error came about. Modify the _extractVariables()_ function to fix the problem.

```{r}

extractVariables =
   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
   {
      # Find the index of the row with =s
      eqIndex <- grep("^===", file)
      spacerRow <- file[eqIndex]
      headerRow <- tolower(file[ eqIndex - 1 ])
      body <- file[ -(1 : eqIndex) ]
      
      blank <- grep("^[[:blank:]]*$", body)
      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
      ignore <- union(blank, footnote)
      
      if(length(ignore))
         body <- body[-ignore]
      
      # Obtain the starting and ending positions of variables
      searchLocs <- findColLocs(spacerRow)
      locCols <- selectCols(varNames, headerRow, searchLocs)
      
      Values <- mapply(substr, list(body), start = locCols[1, ],
                       stop = locCols[2, ])
      colnames(Values) <- varNames
      
      invisible(Values)
   }

```

### 5.)

Write the _convertTime()_ function described in Section 2.3. This function takes a string where time is in either the format hh:mm:ss or mm:ss. The return value is the time as numeric value of the number of minutes. Design this function to take a character vector with multiple strings and return a numeric vector.

```{r}

convertTime <- function( charTime ) {

   timePieces <- strsplit(charTime, ":")

   timePieces <- sapply(timePieces, as.numeric)

   runTime <- sapply(timePieces,
                  function(x) {
                     if(length(x) == 2) x[1] + x[2]/60
                     else 60 * x[1] + x[2] + x[3]/60
                  })
   
}

```

### 6.)

Modify the _createDF()_ function in Section 2.3 to handle the formatting problem with the 2006 male file. You will need to carefully inspect the raw text file in order to determien the problem.

```{r}

createDF =
   function(Res, year, sex) {
      useTime <- if(!is.na(Res[1, 'net']) )
                     Res[, 'net']
                  else if( !is.na(Res[1, 'gun']) )
                     Res[ , 'gun']
                  else
                     Res[, 'time']
      
      useTime <- gsub("[#\\*[:blank:]]", "", useTime)
      
      Res <- Res[ useTime != "", ]
      
      runTime <- convertTime(useTime[ useTime != "" ])
      
      N <- nrow(Res)
      
      Results <- data.frame( year = rep(year, N),
                             sex = rep(sex, N),
                             name = Res[, 'name'],
                             home = Res[, 'home'],
                             age = as.numeric(Res[, 'ag']),
                             runTime = runTime,
                             stringsAsFactors = F)
      
      invisible(Results)
   }

```

### 7.)

Follow the approach developed in Section 2.2 to read the files for the female runners and then process them using the functions in Section 2.3 to create a data frame for analysis. You may need to generalize the _createDF()_ and _extractVariables()_ functions to handle additional oddities in the raw text files.

```{r}

womenDF <- mapply(createDF, womenResMat, year = years,
                  sex = rep("F", 14), SIMPLIFY = F)

sapply(womenDF, function(x) sum(is.na(x$runTime)))

```

### 8.)

Modify the call to the _plot()_ function that created figure 2.6 to create Figure 2.7. To do this, read the documentation for plot() to determine which parameters could be helpful.

```{r}
ggplot(cbMen, aes(age, runTime)) +
   geom_jitter(col = Purple8A) +
   geom_smooth() +
   xlim(10, 75) +
   ylim(45, 150)
```

### 9.)

Modify the piecewise linear fit from Section 2.4.2 to include a hing a 70. Examine the coefficients from the fit and compare the fitted curve to the loess curve. Does the additional hing improve the fit?

```{r}
age20to70 <- seq( from = 20, to = 70, by = 1)

mR.lo992 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 1999,])
mR.lo.pr992 <- predict(mR.lo99, data.frame(age = age20to70))

summary(mR.lo.pr992)

mR.lo122 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 2012, ])
mR.lo.pr122 <- predict(mR.lo122, data.frame(age = age20to70))

summary(mR.lo122)

plot(mR.lo.pr992 ~ age20to70,
     type = "l", col = "purple", lwd = 3,
     xlab = "Age (years)", ylab = "Fitted Run Time (minutes)")

lines(x = age20to70, y = mR.lo.pr122,
      col = "green", lty = 2, lwd = 3)

legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
       legend =  c("1999", "2012"), bty = "n")
```

### 10.)

We have seen that the 1999 runners were typically older than the 2012 runners. Compare the age distribution of the runners agross all 14 years of the races. Use quantile-quantile plots, boxplots and density curves to make your comparisions.

```{r}
ggplot(cbMen, aes(age, group = year)) +
   geom_histogram(aes(fill = ..count..), bins = 30) +
   facet_wrap(~year)

ggplot(cbMen, aes(age, group = year)) +
   geom_boxplot(aes(fill = as.factor(year))) +
   coord_flip() +
   labs(title = "Age by Race Year", xlab = "Race Year", ylab = "Age")

ggplot(cbMen, aes(year, age)) +
   geom_quantile()
```

### 11.)

Normalize each male runner's time by the fastest time for the runner of the same age. To do this, find the fastest runnerk for each year of age from 20 to 80, tapply() function maybe helpful here. Smooth these times using loess() and find the smoothed fime using predict. Use these smoothed times to normalize each run time.

```{r}

cbMenSub <- as.data.table(cbMenSub)

fastest_times <- cbMenSub[, .(max = max(runTime)), by = list(year, age)]

normalized_times <- cbMenSub[, .(runTime), by = list(year, age)]
normalized_times <- merge(normalized_times, fastest_times, by = c("year", "age"))
normalized_times[, normTime := runTime/max]

mR.norm <- loess(normTime ~ age, data = normalized_times)

summary(mR.norm)

age_range <- seq( from = 20, to = 80, by = 1 )

smoothed_times <- predict(mR.norm, newdata = age_range)

smooth_results <- data.table(age = age_range, time = smoothed_times)

ggplot(smooth_results, aes(age, time)) +
   geom_point(aes(col = age)) +
   geom_smooth(alpha = .2)
```

### 12.)

Clean the strings in home and menRes to remove all leading and trailing blanks and multiple contiguous blanks. Also, make all letters lower case and remove any punctuation such as '.' or ','.

```{r}

homeClean2 <- str_remove_all(homeClean, " ")
homeClean2 <- str_remove_all(homeClean2, ",")

head(homeClean2)
```

### 13.)

In section 2.5 we created an id for a runner by pasting together name, year of birth, and state. Consider using the home town instead of the state. How many runners have competed in at least 8 races using this new id?

```{r}
cbMenSub2 <- cbMenSub

cbMenSub2$ID = paste(nameClean, homeClean2, cbMenSub$yob, sep = "_")

races <- tapply(cbMenSub2$year, cbMenSub2$ID, length)

races8 <- names(races)[which(races >= 8)]

men8 <- cbMenSub2[ cbMenSub2$ID %in% races8, ]

orderByRunner <- order(men8$ID, men8$year)

men8 <- men8[orderByRunner, ]
```

### 14.)

Further refine the set of athletes in the longitudinal analysis by dropping those IDs who have a large jump in time in consecutive races and who did not compete for two or more years in a row. How many unique IDs do you have when you include these additional restrictions?

```{r}
numRaces <- tapply(cbMenSub$year, cbMenSub$ID, length)
races8 <- names(numRaces)[which(numRaces >= 8)]
men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]

by_year <- as.data.table(men8)[, .(year, runTime), by = ID]

sans_outliers <- by_year %>%
   group_by(ID) %>%
   mutate(year_gap = year - lag(year, 1), run_diff = runTime - lag(runTime, 1)) %>%
   filter(year_gap <= 2) %>%
   filter(abs(run_diff) < 5) %>%
   distinct(ID)

nrow(sans_outliers)
```
